---
title: "inital-exploration"
output: html_document
---

```{r, message = FALSE, warning = FALSE}
## load necessary packages

library(tidyverse)
library(plotly)
library(lubridate)
```

```{r, warning = FALSE, message = FALSE}
## read CSV into R

firewall_0406_original <- read_csv("Firewall-04062012.csv") 
```

```{r}
## rename columns and parse date/time to aid analysis

firewall_0406 <- firewall_0406_original %>%
  mutate(date_time = dmy_hms(`Date/time`)) %>%
  rename(syslog_priority = `Syslog priority`) %>%
  rename(operation = Operation) %>%
  rename(message_code = `Message code`) %>%
  rename(protocol = Protocol) %>%
  rename(source_IP = `Source IP`) %>%
  rename(destination_IP = `Destination IP`) %>%
  rename(source_hostname = `Source hostname`) %>%
  rename(destination_hostname = `Destination hostname`) %>%
  rename(source_port = `Source port`) %>%
  rename(destination_port = `Destination port`) %>%
  rename(destination_service = `Destination service`) %>%
  rename(direction = Direction) %>%
  rename(connections_built = `Connections built`) %>%
  rename(connections_torn_down = `Connections torn down`)
```

# April 6 Data

## Group by source IP address and time, summarize to get volume of communication 

First, I grouped the data by hour and source_IP with the intention of plotting a line chart with time on the x axis and number of communications from the source on the y axis. The goal was to look for unusual spikes in activity. 

```{r, message = FALSE, warning = FALSE}
## grouping data by source and hour to look for spikes and patterns in the data; defining a new variable num_comms_from_IP that counts how many times the source computer sent a communcation

firewall_0406 %>%
  group_by(source_IP, hour(date_time)) %>%
  summarize(
    num_comms_from_IP = n()
  ) %>%
  rename(time = "hour(date_time)") %>%
  arrange(desc(num_comms_from_IP))
```

Unfortunately, however, there seem to be far too many source IP addresses to plot each one individually on a line chart as I may have envisoned. As we can see below, there are 2990 different source IPs represented in the data set.

```{r}
## grouping data just by source IP to see how many source IP addresses there are over the data set; calculated number of communications initiated by that machine to identify outliers and sorted by that number of communications (descending) to identify possible machines that would be "spamming" otheer computers

firewall_0406 %>%
  group_by(source_IP) %>%
  summarize(
    num_comms_from_IP = n()
  ) %>%
  arrange(desc(num_comms_from_IP))
```
However, one computer -- 172.23.0.132 -- seems to be "spamming" everyone else. This machine is sending almost nine times as many communications as the next most frequently communicating machine.    

```{r, message = FALSE, warning = FALSE}
fireewall_0406_suspicious_IP_for_graph <- firewall_0406 %>%
  filter(source_IP %in% c("172.23.0.132", "172.23.233.58", "172.23.233.50",  "172.23.232.255", "172.23.233.59", "172.23.233.53", "172.23.233.51", "172.23.235.58")) %>%
  group_by(hour(date_time), source_IP) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  rename(time = "hour(date_time)")
```

```{r}
suspicious_IP_comms <- ggplot(fireewall_0406_suspicious_IP_for_graph, aes(x = time, y = num_comms_sent, color = source_IP)) +
  geom_line() + 
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(name = "Time on 04/06 (by hour)") +
  scale_y_continuous(name = "Number of Communications Sent") +
  ggtitle("Aggregated communications sent by 172.23.0.132 on 04/06") +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black")) +
  
  ## -------------------------------------------------------
  
  ## Annotate idea and syntax from http://www.sthda.com/english/wiki/ggplot2-texts-add-text-annotations-to-a-graph-in-r-software
  
  annotate(geom = "text", x = 6, y = 1200, label = "Below this text:", color = "black", size = 2.5) +
  annotate(geom = "text", x = 6, y = 1000, label = "next 7 most frequently", color = "black", size = 2.5) +
  annotate(geom = "text", x = 6, y = 800, label = "communicating IP address", color = "black", size = 2.5)

## ------------------------------------------------------
  

ggplotly(suspicious_IP_comms)
```

While looking at the communications by time does not necessarily elucdiate any further patterns, we can look at to whom this IP address is communicating.

```{r}
suspicious_IP_by_dest <- firewall_0406 %>%
  filter(source_IP == "172.23.0.132") %>%
  group_by(source_IP, destination_IP) %>%
  summarize(
    num_comms_sent = n()
  )

ggplot(suspicious_IP_by_dest, aes(x = num_comms_sent, y = destination_IP)) +
  geom_col()
```

The majority of these communications seem to be going to the IP address 10.32.0.100, which is the "Corporate Headquarters DataCenter firewall interface to the Internet." The rest of the IP addresses are websites on the internet. 

```{r, message = FALSE, warning = FALSE}
suspicious_IP_by_dest_and_time <- firewall_0406 %>%
  filter(source_IP == "172.23.0.132") %>%
  group_by(source_IP, destination_IP, hour(date_time)) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  rename(time = "hour(date_time)")

suspicious_IP_comms_by_dest <- ggplot(suspicious_IP_by_dest_and_time, aes(x = time, y = num_comms_sent, color = destination_IP)) +
  geom_line() +
   ## palette idea from: https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/
  scale_color_brewer(palette = "Set3") +
  scale_x_continuous(name = "Time on 04/06 (by hour)") +
  scale_y_continuous(name = "Number of Communications Sent") +
  ggtitle("Communications sent by IP address 172.23.0.132 on 04/06") +
  ## labs syntax from https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
  labs(color = "Destination IP Address") +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black"))

ggplotly(suspicious_IP_comms_by_dest)
```
Communications from 172.23.0.32 to 10.32.0.100 have a very different pattern and frequency when compared to communications to other IP addresses. 

```{r}
firewall_0406 %>%
  filter(source_IP == "172.23.0.132") %>%
  group_by(destination_IP, destination_port) %>%
  summarize(
    num_comms = n()
  ) %>%
  arrange(desc(num_comms))
```

## Group by destination IP address and time, summarize to get volume of communication 

```{r, message = FALSE, warning = FALSE}
firewall_0406 %>%
  group_by(destination_IP, hour(date_time)) %>%
  summarize(
    n = n()
  ) %>%
  arrange(desc(n))
```
# April 7

```{r, warning = FALSE, message = FALSE}
## read CSV into R

firewall_0407_original <- read_csv("Firewall-04072012.csv") 
```

```{r}
## rename columns and parse date/time to aid analysis

firewall_0407 <- firewall_0407_original %>%
  mutate(date_time = dmy_hms(`Date/time`)) %>%
  rename(syslog_priority = `Syslog priority`) %>%
  rename(operation = Operation) %>%
  rename(message_code = `Message code`) %>%
  rename(protocol = Protocol) %>%
  rename(source_IP = `Source IP`) %>%
  rename(destination_IP = `Destination IP`) %>%
  rename(source_hostname = `Source hostname`) %>%
  rename(destination_hostname = `Destination hostname`) %>%
  rename(source_port = `Source port`) %>%
  rename(destination_port = `Destination port`) %>%
  rename(destination_service = `Destination service`) %>%
  rename(direction = Direction) %>%
  rename(connections_built = `Connections built`) %>%
  rename(connections_torn_down = `Connections torn down`)
```

```{r}
firewall_0407 %>%
  group_by(source_IP, hour(date_time)) %>%
  summarize(
    num_comms_from_IP = n()
  ) %>%
  rename(time = "hour(date_time)") %>%
  arrange(desc(num_comms_from_IP))
```

```{r}
firewall_0407 %>%
  group_by(source_IP)%>%
  summarize(
    num_comms_from_IP = n()
  ) %>%
  arrange(desc(num_comms_from_IP))
```
While 172.23.0.132, the identified suspicious machine from the April 6th data, remains a frequently communicating source here, another individual workstation computer, 172.23.252.10 

```{r, message = FALSE, warning = FALSE}
firewall_0406 %>%
  filter(source_IP == "172.23.252.10")
```

This computer is not found in the 4/6 data.

I'm going to look at this data by time

```{r, message = FALSE, warning = FALSE}
fireewall_0407_suspicious_IP_for_graph <- firewall_0407 %>%
  filter(source_IP %in% c("172.23.252.10", "172.23.0.132", "172.23.232.253",  "172.23.232.252", "172.23.232.9", "172.23.233.36", "172.23.235.53", "172.23.233.33")) %>%
  group_by(hour(date_time), source_IP) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  rename(time = "hour(date_time)")
```

```{r}
suspicious_IP_comms07 <- ggplot(fireewall_0407_suspicious_IP_for_graph, aes(x = time, y = num_comms_sent, color = source_IP)) +
  geom_line() + 
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(name = "Time on 04/07 (by hour)") +
  scale_y_continuous(name = "Number of Communications Sent") +
  ggtitle("172.23.0.132 and 172.23.252.10 on 04/07") +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black")) +
  labs(caption = "Next 6 most frequently communicating IP addresses added for reference")
  

ggplotly(suspicious_IP_comms07)
```
```{r, message = FALSE, warning = FALSE}
suspicious_IP_by_dest_and_time132 <- firewall_0407 %>%
  filter(source_IP == "172.23.0.132") %>%
  group_by(source_IP, destination_IP, hour(date_time)) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  rename(time = "hour(date_time)")

suspicious_IP_comms_by_dest132 <- ggplot(suspicious_IP_by_dest_and_time132, aes(x = time, y = num_comms_sent, color = destination_IP)) +
  geom_line() +
   ## palette idea from: https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/
  scale_color_brewer(palette = "Set3") +
  scale_x_continuous(name = "Time on 04/07 (by hour)") +
  scale_y_continuous(name = "Number of Communications Sent") +
  ggtitle("Communications sent by 172.23.0.132 on 04/07") +
  ## labs syntax from https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
  labs(color = "Destination IP Address") +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black"))

ggplotly(suspicious_IP_comms_by_dest132)
```


```{r}
firewall_0407 %>%
  filter(source_IP == "172.23.0.132") %>%
  group_by(destination_IP, destination_port) %>%
  summarize(
    num_comms = n()
  ) %>%
  arrange(desc(num_comms))
```


```{r, message = FALSE, warning = FALSE}
suspicious_IP_by_dest_and_time10 <- firewall_0407 %>%
  filter(source_IP == "172.23.252.10") %>%
  group_by(source_IP, destination_IP, hour(date_time)) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  rename(time = "hour(date_time)")

suspicious_IP_comms_by_dest10 <- ggplot(suspicious_IP_by_dest_and_time10, aes(x = time, y = num_comms_sent, color = destination_IP)) +
  geom_line() +
   ## palette idea from: https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/
  scale_color_brewer(palette = "Set3") +
  scale_x_continuous(name = "Time on 04/07 (by hour)") +
  scale_y_continuous(name = "Number of Communications Sent") +
  ggtitle("Communications sent by 172.23.252.10 on 04/07") +
  ## labs syntax from https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
  labs(color = "Destination IP Address") +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black"))

ggplotly(suspicious_IP_comms_by_dest10)
```
```{r}
suspicious_IP_by_dest_and_time10p2 <- firewall_0407 %>%
  filter(source_IP == "172.23.252.10") %>%
  group_by(source_IP, destination_IP, hour(date_time), destination_port) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  rename(time = "hour(date_time)")

suspicious_IP_comms_by_dest10p2 <- ggplot(suspicious_IP_by_dest_and_time10p2, aes(x = time, y = num_comms_sent, color = destination_port)) +
  geom_line() +
   ## palette idea from: https://www.datanovia.com/en/blog/the-a-z-of-rcolorbrewer-palette/
  scale_color_brewer(palette = "Set2") +
  scale_x_continuous(name = "Time on 04/07 (by hour)") +
  scale_y_continuous(name = "Number of Communications Sent") +
  ggtitle("Communications sent by 172.23.252.10 on 04/07") +
  ## labs syntax from https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
  labs(color = "Destination \n Port") +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black"))

ggplotly(suspicious_IP_comms_by_dest10p2)
```



```{r}
firewall_0407 %>%
  filter(source_IP == "172.23.252.10") %>%
  group_by(destination_IP, destination_port) %>%
  summarize(
    num_comms = n()
  ) %>%
  arrange(desc(num_comms))
```

```{r}
firewall_0407 %>%
  filter(source_IP == "172.23.0.132") %>%
  group_by(destination_IP, destination_port) %>%
  summarize(
    num_comms = n()
  ) %>%
  arrange(desc(num_comms))
```

## Group by destination IP address and time, summarize to get volume of communication 

```{r, message = FALSE, warning = FALSE}
firewall_0407 %>%
  group_by(destination_IP) %>%
  summarize(
    n = n()
  ) %>%
  arrange(desc(n))
```
This data set shows a similar pattern to the data from 4/6; 10.32.0.100 has by far the most frequent communications, followed by other IP addresses in 10.32.x.xx.



```{r, message = FALSE, warning = FALSE}
dest_port1 <- firewall_0406 %>%
  filter(destination_port == 6667) %>%
  group_by(source_IP) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  arrange(desc(num_comms_sent))

dest_port1_by_time <- firewall_0406 %>%
  filter(destination_port == 6667) %>%
  group_by(source_IP, hour(date_time)) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  arrange(desc(num_comms_sent))

dest_port2 <- firewall_0407 %>%
  filter(destination_port == 6667) %>%
  group_by(source_IP) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  arrange(desc(num_comms_sent)) %>%
  filter(num_comms_sent > 4626)

dest_port2_by_time <- firewall_0407 %>%
  filter(destination_port == 6667) %>%
  group_by(source_IP, hour(date_time)) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  arrange(desc(num_comms_sent))

ggplot(dest_port2, aes(x = num_comms_sent, y = source_IP)) +
  geom_col() +
  scale_x_continuous(name = "Number of Communications to Destination Port 6667", limits = c(0,50000)) +
  scale_y_discrete(name = "Source IP (top 10 with comms to destination port 6667)") +
  ggtitle("Top Source IPs with Communications to Dest. Port 6667 on 4/7") +
  ## labs syntax from https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black"))

  
```

```{r}
dest_port2_by_time_twoIP <- firewall_0407 %>%
  filter(destination_port == 6667) %>%
  filter(source_IP == "172.23.252.10" | source_IP == "172.23.0.108" | source_IP == "172.23.1.168" | source_IP == "172.23.134.76" | source_IP == "172.23.134.214" | source_IP == "172.23.134.218" | source_IP == "172.23.134.78") %>%
  group_by(source_IP, hour(date_time)) %>%
  summarize(
    num_comms_sent = n()
  ) %>%
  arrange(desc(num_comms_sent)) %>%
  rename(time = "hour(date_time)")

dest_port_by_IP2 <- ggplot(dest_port2_by_time_twoIP, aes(x = time, y = num_comms_sent, color = source_IP)) +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(name = "Time on 04/07 (by hour)") +
  scale_y_continuous(name = "Number of Communications Sent to Port 6667") +
  ggtitle("Communications sent to Dest. Port 6667 on 04/07") +
  ## labs syntax from https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
  labs(color = "Source IP") +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black")) +
   ## Annotate idea and syntax from http://www.sthda.com/english/wiki/ggplot2-texts-add-text-annotations-to-a-graph-in-r-software
  
  annotate(geom = "text", x = 6, y = 800, label = "Below this text:", color = "black", size = 2.5) +
  annotate(geom = "text", x = 6, y = 700, label = "next 5 source IP addresses with greatest", color = "black", size = 2.5) +
  annotate(geom = "text", x = 6, y = 600, label = "communication in any single hour to port 6667", color = "black", size = 2.5) +
  annotate(geom = "text", x = 6, y = 500, label = "for reference", color = "black", size = 2.5)

## ------------------------------------------------------

ggplotly(dest_port_by_IP2)
```

```{r}
firewall_0406 %>%
  filter(source_IP == "172.23.0.108") %>%
  group_by(destination_port) %>%
  summarize(
    num_comms = n()
  )

firewall_0407 %>%
  filter(source_IP == "172.23.0.108") %>%
  group_by(destination_port) %>%
  summarize(
    num_comms = n()
  )

firewall_0407 %>%
  filter(source_IP == "172.23.0.108") %>%
  group_by(destination_port, destination_IP) %>%
  summarize(
    num_comms = n()
  )

suspicious_IP3 <- firewall_0407 %>%
  filter(source_IP == "172.23.0.108") %>%
  group_by(destination_port, destination_IP, hour(date_time)) %>%
  summarize(
    num_comms = n()
  ) %>%
  rename(time = "hour(date_time)")

plot <- ggplot(suspicious_IP3, aes(x = time, y = num_comms, color = destination_IP)) +
  geom_line() +
  facet_wrap(~destination_port, labeller = "label_both") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(name = "Time on 04/07 (by hour)") +
  scale_y_continuous(name = "Number of Communications Sent to Destination Port") +
  ggtitle("Communications sent by 172.23.0.108 04/07") +
  ## labs syntax from https://www.datanovia.com/en/blog/ggplot-legend-title-position-and-labels/
  labs(color = "Destination IP") +
  theme(panel.grid = element_blank(), panel.background = element_rect(fill = "white"), axis.line = element_line(color = "black"))
  

ggplotly(plot)
```


```{r}
firewall_0407 %>%
  filter(source_IP == "172.23.252.10") %>%
  group_by(hour(date_time), minute(date_time)) %>%
  arrange(date_time)
```












